---
title: "cbs_workflow"
author: "Nazia Jassim"
date: "04/10/2020"
output: html_document
---
## Set wd
```{r}
getwd();
library(tidyverse);
library(BayesFactor);
library(lsr);
library(ggforce);
library(ggdist);
library(gghalves);

```

## Data cleaning and prep for analysis
```{r}

## Read in dataset 
cbs <- read_csv("cbs_dataset_v1.csv") %>% as_tibble
## Remove spaces and punctuations from colnames
names(cbs) <- str_replace_all(names(cbs), c(" " = "." , "," = "" ));
## Select variables of interest
cbs <- select(cbs, Subject_ID, Digit.Span_Attempted: Rotations_Task.Duration, cohort.type, cohort.phase, AQ, Autism.diagnosis, Gender, Age, Highest.education, SPQ, Relatives.with.autism, synaesthesia)
## prep groups- sex and diagnosis
 cbs %>% 
## Only Male and Female sexes selected (sorry non-binary participants:(!)
   filter(Gender== "Male" | Gender== "Female") %>%
   mutate(cohort = as_factor(cohort.type), sex = as_factor(Gender), AQ= as.numeric(AQ)) -> cbs_clean
## create separate male and female groups for use later
cbs_males <- filter(cbs_clean, sex =="Male")
cbs_females <- filter( cbs_clean, sex== "Female")
```


## Create cutoffs based on working memory (WM) perfomance
## Participants whose WM performance falls below cutoff removed from further analysis
## Cutoffs computed based on overall group mean
## !!!  CHANGE DATASET NAME IF REQUIRED!
```{r}

# digit span and monkey ladders
 cbs_clean %>% 
  #group_by(cohort) %>%  
  mutate(cohort= ifelse(cohort=="asc", "Autism", "Control"), ds_cutoff= mean(Digit.Span_Single.Score.Raw) - 2*sd(Digit.Span_Single.Score.Raw),   ml_cutoff=mean(Monkey.Ladder_Single.Score.Raw) - 2*sd(Monkey.Ladder_Single.Score.Raw)) %>%
filter(Digit.Span_Single.Score.Raw > ds_cutoff & Monkey.Ladder_Single.Score.Raw > ml_cutoff) %>% 
select(-ds_cutoff & -ml_cutoff) %>% as_data_frame() -> cbs_wm 

# t tests to check if groups differ in working mermory
# Digit span
t.test(Digit.Span_Single.Score.Raw ~ cohort, cbs_wm);
ttestBF(formula=Digit.Span_Single.Score.Raw ~cohort, data=cbs_wm);
cohensD(Digit.Span_Single.Score.Raw ~cohort, cbs_wm);

# summary digit span
cbs_wm%>% 
  group_by(cohort)%>%
  summarise(mean= mean(Digit.Span_Single.Score.Raw), sd= sd(Digit.Span_Single.Score.Raw))
                                        

## Digit span plots
cbs_wm %>% ggplot(mapping= aes(x=cohort, y= Digit.Span_Single.Score.Raw,fill= cohort)) + xlab("Group")+ ylab("Digit Span Score")+ labs(fill= "Group", title = "S1: Working Memory- Digit Span Scores")+
    ggdist::stat_halfeye(adjust = .5, width = .7, .width = 0, justification = -.2, point_colour = NA) + 
    geom_boxplot(width = .2, outlier.shape = NA) + 
    geom_jitter(width = .08, size= 1, alpha= 0.3)


                                        
# Monkey ladders
t.test(Monkey.Ladder_Single.Score.Raw~cohort, cbs_wm);
ttestBF(formula=Monkey.Ladder_Single.Score.Raw~cohort, data=cbs_wm);
cohensD(Monkey.Ladder_Single.Score.Raw~cohort, cbs_wm);

# summary 
cbs_wm%>% 
  group_by(cohort)%>%
  summarise(mean= mean(Monkey.Ladder_Single.Score.Raw), sd= sd(Monkey.Ladder_Single.Score.Raw))

# Boxplot with raincloud
cbs_wm %>% ggplot(mapping= aes(x=cohort, y= Monkey.Ladder_Single.Score.Raw,fill= cohort)) + xlab("Group")+ ylab("Monkey Ladder Score")+ labs(fill= "Group", title = "S2: Working Memory- Monkey Ladder Scores")+
    ggdist::stat_halfeye(adjust = .5, width = .7, .width = 0, justification = -.2, point_colour = NA) + 
    geom_boxplot(width = .2, outlier.shape = NA) + 
    geom_jitter(width = .08, size= 1, alpha= 0.3)
```

## Summary stats for final dataset to be used for further analysis
```{r}

length(cbs_wm$cohort =="asc");
sum(cbs_wm$cohort =="asc" &cbs_wm$sex== "Female");
sum(cbs_wm$cohort =="asc" &cbs_wm$sex== "Male");
length(cbs_wm$cohort =="ctr");
sum(cbs_wm$cohort =="ctr" &cbs_wm$sex== "Male");
sum(cbs_wm$cohort =="ctr" &cbs_wm$sex== "Female");

# age
cbs_wm%>% 
group_by(cohort)%>%
summarise(mean= mean(Age, na.rm=TRUE), 
            median= median(Age, na.rm=TRUE), 
            sd_age= sd(Age, na.rm=TRUE), 
            iqr= IQR(Age, na.rm=TRUE),
            max= max(Age),
            min= min(Age))

```

## Task 1- Feature Match
```{r}
##### I CBS Scores ---------------------------------------------------
## Remove outliers, computed separately for each diagnosis group
data_clean_FM <- cbs_wm %>%  
  group_by(cohort) %>%  
  mutate(cutoff= mean(Feature.Match_Score) - 2*sd(Feature.Match_Score)) %>%   filter(Feature.Match_Score > cutoff) %>% 
select(-cutoff) 

## Summary stats 
data_clean_FM %>%
  group_by(cohort) %>%
  summarise(mean= mean(Feature.Match_Score, na.rm=TRUE), 
            median= median(Feature.Match_Score, na.rm=TRUE), 
            sd= sd(Feature.Match_Score, na.rm=TRUE), 
            iqr= IQR(Feature.Match_Score, na.rm=TRUE),
            max= max(Feature.Match_Score),
            min= min(Feature.Match_Score));
## Plots 
ggplot(data_clean_FM, mapping= aes(x=cohort, y= Feature.Match_Score, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9) + facet_wrap(~sex, ncol=2);

ggplot(data_clean_FM, mapping= aes(x=AQ,na.rm=TRUE, y= Feature.Match_Score, color= cohort))+ geom_point();

## Distribution
 ggplot(data = data_clean_FM) +
 geom_density(mapping = aes(x = Feature.Match_Score, fill = cohort, colour= cohort), alpha= 0.2) + facet_wrap(~sex, ncol=2)
 
```

```{r}
#----------------------------------------------------------------------------#
 
#------------------------------- II ACCURACY RATES  --------------------------
 
 # Overall accuracy- 

## Compute accuracy scores and summary stats
#   fm_accuracy <- cbs_wm %>%
#   mutate(fm_accuracy_score = (Feature.Match_Attempted-Feature.Match_Errors)) 
#   
#   fm_accuracy %>%
#   group_by(cohort) %>%
#   summarise(mean = mean(fm_accuracy_score),
#             median= median(fm_accuracy_score),
#             sd = sd(fm_accuracy_score),
#             max= max(fm_accuracy_score),
#             min= min(fm_accuracy_score))
#   
# ## Remove outliers for accuracy scores (ie less than 2SD of group mean)
#   fm_accuracy_clean <- fm_accuracy %>%
#   group_by(cohort) %>%  
#   mutate(cutoff= mean(fm_accuracy_score)- 2*sd(fm_accuracy_score)) %>% 
#         filter( fm_accuracy_score > cutoff) %>% 
#         select(-cutoff)
  
  # ------------------------------------------------ # --------------------------------------------------------------------------------#
  
## Compute Accuracy Rates 
cbs_wm %>% as_tibble() %>% select("Subject_ID", "AQ", "Age", "sex", "Gender","cohort.type", c(12:26),"Subject_ID") %>%
   mutate(cohort = as.factor(cohort.type),
     fm_accuracy_score = (Feature.Match_Attempted-Feature.Match_Errors),
          fm_accuracy_overall = (fm_accuracy_score/Feature.Match_Attempted*100), # overall accuracy
          fm_match_errors = (Feature.Match_Match.Attempted - Feature.Match_Match.Correct),
          fm_accuracy_match = ((Feature.Match_Match.Attempted - fm_match_errors)/(Feature.Match_Match.Attempted)*100),  # match
          fm_mismatch_errors = (Feature.Match_Mismatch.Attempted - Feature.Match_Mismatch.Correct),
          fm_accuracy_mismatch = (Feature.Match_Mismatch.Attempted - fm_mismatch_errors)/(Feature.Match_Mismatch.Attempted)*100) ->  fm_accuracy;


# summary stats
fm_accuracy %>%
   group_by(cohort) %>%
  summarise(mean = mean(fm_accuracy_overall),
            median= median(fm_accuracy_overall),
            sd = sd(fm_accuracy_overall),
             max= max(fm_accuracy_overall),
             min= min(fm_accuracy_overall))


# remove outliers by overall accuracy 
fm_accuracy %>%
group_by(cohort) %>%  
mutate(cutoff= mean(fm_accuracy_overall)- 2*sd(fm_accuracy_overall)) %>% 
filter(fm_accuracy_overall> cutoff) %>%  
  select(-cutoff) ->fm_accuracy_clean

fm_accuracy_clean %>%
   group_by(cohort) %>%
  summarise(mean = mean(fm_accuracy_overall),
            median= median(fm_accuracy_overall),
            sd = sd(fm_accuracy_overall),
             max= max(fm_accuracy_overall),
             min= min(fm_accuracy_overall))

# arrange acc by condition 
fm_accuracy %>%
  pivot_longer(cols=c('fm_accuracy_match','fm_accuracy_mismatch'), names_to= "condition", values_to= "accuracy_rate_per_condition") %>% 
  mutate(condition=as.factor(condition)) -> fm_accuracy_by_condition;

fm_accuracy_by_condition$condition = ifelse(fm_accuracy_by_condition$condition=="fm_accuracy_match", "match", "mismatch");

## PLOT
 # overall
ggplot(fm_accuracy_by_condition, mapping= aes(x=cohort, y= fm_accuracy_overall, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9) + facet_wrap(~condition)

## Write to csv 
write.csv(fm_accuracy_by_condition, "fm_accuracy.csv")
```

## Task 2- Polygons
```{r}

## Remove outliers
data_clean_poly <- cbs_wm %>%  
  group_by(cohort) %>%  
  mutate(cutoff= mean(Polygons_Score) - 2*sd(Polygons_Score)) %>%   filter(Polygons_Score > cutoff) %>% 
select(-cutoff) 

## Summarize scores
data_clean_poly %>%
  group_by(cohort) %>%
  summarise(mean= mean(Polygons_Score, na.rm=TRUE), 
            median= median(Polygons_Score, na.rm=TRUE), 
            sd= sd(Polygons_Score, na.rm=TRUE), 
            iqr= IQR(Polygons_Score, na.rm=TRUE),
            max= max(Polygons_Score),
            min= min(Polygons_Score))
## Plots
ggplot(data_clean_poly, mapping= aes(x=cohort, y= Polygons_Score, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9) + facet_wrap(~sex, ncol=2);

ggplot(data_clean_poly, mapping= aes(x=AQ,na.rm=TRUE, y= Polygons_Score, color= cohort))+ geom_point()

## Distribution
ggplot(data = data_clean_poly) +
geom_density(mapping = aes(x = Polygons_Score, fill = cohort, colour= cohort), alpha= 0.2) + facet_wrap(~sex, ncol=2)
```

```{r}
#----------------------------------------------------------------------------#
 
#------------------------------- II POLYGONS ACCURACY RATES  --------------------------
 
 # Overall accuracy- 

## Compute Accuracy Rates 
cbs_wm %>% as_tibble() %>% select("Subject_ID", "AQ", "Age", "sex", "Gender","cohort.type", c(38:50),"Subject_ID") %>%
   mutate(cohort = as.factor(cohort.type),
     poly_accuracy_score = (Polygons_Attempted-Polygons_Errors),
          poly_accuracy_overall = (poly_accuracy_score/Polygons_Attempted*100), # overall accuracy
          poly_match_errors = (Polygons_Match.Attempted - Polygons_Match.Correct),
          poly_accuracy_match = ((Polygons_Match.Attempted - poly_match_errors)/(Polygons_Match.Attempted)*100),  # match
          poly_mismatch_errors = (Polygons_Mismatch.Attempted - Polygons_Mismatch.Correct),
          poly_accuracy_mismatch = (Polygons_Mismatch.Attempted - poly_mismatch_errors)/(Polygons_Mismatch.Attempted)*100) ->  poly_accuracy;


# summary stats
poly_accuracy %>%
   group_by(cohort) %>%
  summarise(mean = mean(poly_accuracy_overall),
            median= median(poly_accuracy_overall),
            sd = sd(poly_accuracy_overall),
             max= max(poly_accuracy_overall),
             min= min(poly_accuracy_overall))


# remove outliers by overall accuracy 
# poly_accuracy %>%
# group_by(cohort) %>%  
# mutate(cutoff= mean(poly_accuracy_overall)- 2*sd(poly_accuracy_overall)) %>% 
# filter(poly_accuracy_overall> cutoff) %>%  
#   select(-cutoff) ->poly_accuracy_clean
# 
# poly_accuracy_clean %>%
#    group_by(cohort) %>%
#   summarise(mean = mean(poly_accuracy_overall),
#             median= median(poly_accuracy_overall),
#             sd = sd(poly_accuracy_overall),
#              max= max(poly_accuracy_overall),
#              min= min(poly_accuracy_overall))

# arrange acc by condition 
poly_accuracy %>%
  pivot_longer(cols=c('poly_accuracy_match','poly_accuracy_mismatch'), names_to= "condition", values_to= "accuracy_rate_per_condition") %>% 
  mutate(condition=as.factor(condition)) -> poly_accuracy_by_condition;

poly_accuracy_by_condition$condition = ifelse(poly_accuracy_by_condition$condition=="poly_accuracy_match", "match", "mismatch");

poly_accuracy_by_condition  %>% group_by(cohort, condition) %>%summarise(mean= mean(accuracy_rate_per_condition), sd= sd(accuracy_rate_per_condition), m)

## PLOT
 # overall
ggplot(poly_accuracy_by_condition, mapping= aes(x=cohort, y= poly_accuracy_overall, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9) + facet_wrap(~condition)

## Write to csv 
write.csv(poly_accuracy_by_condition, "poly_accuracy.csv")

```

## Task 3- Rotations
```{r}
## Remove outliers
data_clean_rot <- cbs_wm %>%  
  group_by(cohort) %>%  
  mutate(cutoff= mean(Rotations_Score) - 2*sd(Rotations_Score)) %>%   filter(Rotations_Score > cutoff) %>% 
select(-cutoff);
## Summarize 
data_clean_rot %>%
  group_by(cohort) %>%
  summarise(mean= mean(Rotations_Score, na.rm=TRUE), 
            median= median(Rotations_Score, na.rm=TRUE), 
            sd= sd(Rotations_Score, na.rm=TRUE), 
            iqr= IQR(Rotations_Score, na.rm=TRUE),
            max= max(Rotations_Score),
            min= min(Rotations_Score))

## Plot
ggplot(data_clean_rot, mapping= aes(x=cohort, y= Rotations_Score, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9)+ facet_wrap(~sex, ncol=2)

ggplot(data_clean_rot, mapping= aes(x=AQ,na.rm=TRUE, y= Rotations_Score, color= cohort))+ geom_point();

## Distributions
 ggplot(data = data_clean_rot) +
 geom_density(mapping = aes(x = Rotations_Score, fill = cohort, colour= cohort), alpha= 0.2) + facet_wrap(~sex, ncol=2)
```

```{r}
# Overall accuracy- 

## Compute accuracy scores and summary stats
  rot_accuracy <- cbs_wm %>%
  mutate(rot_accuracy_score = (Rotations_Attempted-Rotations_Errors)) 
  
  rot_accuracy %>%
  group_by(cohort) %>%
  summarise(mean = mean(rot_accuracy_score),
            median= median(rot_accuracy_score),
            sd = sd(rot_accuracy_score),
            max= max(rot_accuracy_score),
            min= min(rot_accuracy_score))
  
## Remove outliers for accuracy scores (ie less than 2SD of group mean)
  rot_accuracy_clean <- rot_accuracy %>%
  group_by(cohort) %>%  
  mutate(cutoff= mean(rot_accuracy_score)- 2*sd(rot_accuracy_score)) %>% 
        filter( rot_accuracy_score > cutoff) %>% 
        select(-cutoff)
  
## Compute Accuracy Rates (after outlier removal for accuracy scores)
 rot_accuracy_clean %>% 
   mutate(rot_accuracy_rate = rot_accuracy_score/Rotations_Attempted*100) -> rot_accuracy_clean
  
 rot_accuracy_clean %>%
    group_by(cohort) %>%
    summarise(mean = mean(rot_accuracy_rate),
            median= median(rot_accuracy_rate),
            sd = sd(rot_accuracy_rate),
            max= max(rot_accuracy_rate),
            min= min(rot_accuracy_rate)) 
  
## PLOT
ggplot(rot_accuracy_clean, mapping= aes(x=cohort, y= rot_accuracy_rate, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9)

### Accurary rates for each condition (from outlier-removed data)

## MATCH -------------------------------
 
#Accuracy
rot_match_accuracy <- rot_accuracy_clean %>% 
mutate(rot_match_errors = Rotations_Match.Attempted - Rotations_Match.Correct)%>%
mutate(rot_accuracy_match = (Rotations_Match.Attempted - rot_match_errors)
   /(Rotations_Match.Attempted)*100)   
 
#Summary stats
rot_match_accuracy %>%
group_by(cohort) %>%
  summarise(mean = mean(rot_accuracy_match),
            median= median(rot_accuracy_match),
            sd = sd(rot_accuracy_match),
            max= max(rot_accuracy_match),
            min= min(rot_accuracy_match))

## PLOT
#Box
ggplot(rot_match_accuracy, mapping= aes(x=cohort, y= rot_accuracy_match, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9);

#Density
ggplot(rot_match_accuracy)+
geom_density(mapping = aes(x = rot_accuracy_match, fill = cohort, colour= cohort), alpha= 0.2) 
  

## MISMATCH ------------------------
#Accuracy
rot_mismatch_accuracy <- rot_accuracy_clean %>% 
mutate(rot_mismatch_errors = Rotations_Mismatch.Attempted - Rotations_Mismatch.Correct)%>%
mutate(rot_accuracy_mismatch = (Rotations_Mismatch.Attempted - rot_mismatch_errors)
   /(Rotations_Mismatch.Attempted)*100)   
 
#Summary stats
rot_mismatch_accuracy %>%
group_by(cohort) %>%
  summarise(mean = mean(rot_accuracy_mismatch),
            median= median(rot_accuracy_mismatch),
            sd = sd(rot_accuracy_mismatch),
            max= max(rot_accuracy_mismatch),
            min= min(rot_accuracy_mismatch))

## PLOT
#Box
ggplot(rot_mismatch_accuracy, mapping= aes(x=cohort, y= rot_accuracy_mismatch, fill=cohort))+ geom_boxplot()+ geom_jitter(color="black", size=0.4, alpha=0.9)
#Density
ggplot(rot_match_accuracy)+
geom_density(mapping = aes(x = rot_accuracy_match, fill = cohort, colour= cohort), alpha= 0.2) 


## Write to CSV
write.csv(rot_match_accuracy, "rot_match_accuracy.csv")
write.csv(rot_mismatch_accuracy, "rot_mismatch_accuracy.csv")
```

